Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs

### 2016-11-15 write-ups/PLDI/paper.tex
Add chip info for HD 520
Ideas for Section 4.

@@

\ADComment{Explain carefully that doing this properly would require
  driver support, etc., and then explain what we've actually done.}

@@

\subsection{An Efficient Resizing Barrier}\label{sec:resizingbarrier}

@@

>>>>

We envisage cooperative kernels being implemented as part of the
runtime for a GPU implementation of OpenCL, requiring changes to the
device driver allowing it to intercept our new primitives and schedule
and deschedule workgroups accordingly.

To experiment with our ideas on current GPUs, we have instead
developed a prototype implementation that mocks up the required
runtime support via a \emph{megakernel}, and exploits the
occupancy-bound execution model that these GPUs provide.  To be clear:
an aim of cooperative kernels is to avoid depending on the
occupancy-bound model; our protoytpe exploits this model simply to
allow us to experiment with current GPUs whose proprietary drivers we
cannot change.

\subsection{The megakernel approach}

The mock-up works as follows.  Instead of multitasking multiple
separate kernels, we merge a set of kernels into a megakernel---a
single, monolithic kernel---and encode the scheduling logic inside
this megakernel.  The megakernel is launched with as many workgroups
as can run concurrently according to the occupancy-bound.  One
workgroup takes the role of scheduler, and the remaining workgroups
act as a pool of workers.  A worker repeatedly queries the scheduler
to be assigned a task.  A task corresponds to executing a cooperative
or non-cooperative kernel.  In the case where a task corresponds to a
non-cooperative kernel, the workgroup executes the kernel function
uninterrupted and awaits further work.  In the case of a cooperative
kernel, the workgroup either starts from the kernel entry point,
corresponding to the workgroup being an \emph{initial} workgroup of
the kernel, or immediately jumps to a designated point within the
cooperative kernel, corresponding to the workgroup being forked by an
already executing workgroup.  In the latter case, the new workgroup
also receives a struct containing the values of all relevant
$\transmit$-annotated variables.

\paragraph{Simplifying assumptions}

\ADComment{One cooperative, one non-cooperative.  Justification: sufficient to evaluate the key use case of a long-running irregular algorithm co-existing with a series of short-running tasks.  Simplifieid task of implementing prototype, which we do not want to over-enginneer since it is only a mock-up.}

\ADComment{Restrictions on kernel structure: variables at top of kernel are implicitly $\transmit$-annotated, and $\offerkill$, $\offerfork$ and $\resizingglobalbarrier$ calls must appear only in kernel entry function.  Avoiding $\transmit$ allowed us to reuse the clang OpenCL front-end, and restricitng starting points for fork to be in main kernel allowed a ``goto''-based approach to redirecting control flow; threading this through procedure calls would be possible but laborious.  Restricting kill points enables kill to be implemented as ``return''; exiting from a procedure call stack would require stack unwinding support, akin to throwing an exception.  Again make point that all these issues are only relevant due to our mock-up.}

\paragraph{Merging kernels}

\subsection{Scheduler design}\label{sec:resizingbarrier}

\ADComment{The host and device components of the scheduler, the use of
  SVM, and the host API for scheduling both sorts of kernel.}

@@

@@

\paragraph{An efficient resizing barrier}

@@

\ADComment{Rewrite:} There is scope for combining the approaches by using cooperative scheduling...
